---
title: "MachineLearning_Classification_paudyal"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Loading my Libraries
rm(list = ls())
library(MASS)
library(class)
library(ggplot2)
library(cowplot)
library(caTools)
library(caret)
library(pROC)
library(ROCR)
knitr::opts_chunk$set(echo = TRUE)
```

# Banknote authentication


This dataset presents an example of classification problem (authentic vs. counterfeit bank notes) using continuous 
predictors derived from image processing. More details about underlying data can be found in 
corresponding [dataset description](http://archive.ics.uci.edu/ml/datasets/banknote+authentication) at UCI ML website. 

Loading the data, `data_banknote_authentication.txt` with the following attributes. The first $\textbf{four attributes}$ are used as $\textbf{predictors}$ to classify the outcome (authenic or counterfeit) and is presented by the $\textbf{fifth attribute as "Class"}$: 

1.$\textbf{Var.W}$   : variance of Wavelet Transformed image (continuous)  

2.$\textbf{Skew.W}$  : skewness of Wavelet Transformed image (continuous) 

3.$\textbf{Curt.W}$  : curtosis of Wavelet Transformed image (continuous) 

4.$\textbf{Entropy}$ : entropy of image (continuous) 

5.$\textbf{Class}$   : class (integer) 

```{r Banknote_Authentication}
bank.N <- read.table("data_banknote_authentication.txt",sep=",")
colnames(bank.N) <- c("Var.W","Skew.W","Curt.W","Entropy", "Class")
# Checking the "Predictors" summary
summary(bank.N[,1:4])
# detecting number of ZEROs versus ONEs on the Outcome, "Class" attribute
bank.N$Class <- factor(bank.N$Class)
summary(bank.N[,5])
# The total number of Zeros and Ones are accounted for: 762 + 610 = 1372, all the data has been classified as Authenic or Counterfeit.
dim(bank.N)
```
#Logistic Regression

Fit logistic regression model of the class attribute using the four features as predictors in the model. Based on on the report from the fitting, which features seems to be significantly related to our "Class", a categorical outcome. Use prediction for the entire data (no splitting for train/test). How does the prediction result compare to our "Class" (i.e. confusion Table). 

$\textbf{Logistic Regression Model}$

```{r logistic model}

L.R.fit = glm(formula = Class ~ ., family = "binomial", data = bank.N)
summary(L.R.fit)

```

Summary of the logistic regression is shown above. 


The $\textbf{confusion matrix}$ of our logistic model of the entire dataset is given below

```{r}

predict <- predict(L.R.fit, type = 'response')
#confusion matrix
conf_matrix = table(bank.N[,"Class"], predict > 0.5)
conf_matrix
```
The confusion matrix from the logistic regression gives the good prediction of bank note.

Plotting the link and response of the above prediction.
```{r}
L.R.link <- predict(L.R.fit,type="link")
L.R.response <- predict(L.R.fit,type="response")



class_predict <- data.frame(link=L.R.link, 
                         response=L.R.response,
                         Class = bank.N$Class,
                         stringsAsFactors=FALSE)


  ggplot(class_predict, aes(x=link, y=response, col=Class)) + 
  scale_color_manual(values=c("blue", "red")) + 
  geom_point() + 
  geom_rug() + 
  ggtitle("Classification")
```

Calculate the error rate and explain if this is a training or test error? Also calculate the sensitivity and specificity. 

```{r}
logr_accuracy = sum(diag(conf_matrix))/sum(conf_matrix)
logr_accuracy

logr_error = 1 - logr_accuracy
logr_error

sensitivity = conf_matrix[2,2]/sum(conf_matrix[2,])
sensitivity

specificity = conf_matrix[1,1]/sum(conf_matrix[1,])
specificity

```
The $\textbf{Logistic regression model}$ shows the accuracy of 99.19% and error of 0.80%. The specificity is 99.34% and sensitivity is 99.01%


# LDA and QDA (package: MASS)

Fit LDA and QDA model on the entire dataset and calculate confusion matrix, (training) error rate, sensitivity and specificity for each. Compare them to those of logistic regression. Describe the results.


$\textbf{Quadratic Discriminant Analysis}$
The summary of LDA model is shown below
```{r}
LDA.fit <- lda(Class~., data=bank.N)
summary(LDA.fit)
```

```{r}
LDA.pred <- predict(LDA.fit, bank.N)
c_mat <- table(LDA.pred$class, bank.N$Class)
c_mat

```
Confusion matrix for LDA is shown above. False Negative is really high. It gives us good 1's prediction.

$\textbf{Linear Discriminant Model}$
```{r}
lda_accuracy <- confusionMatrix(c_mat)$overall[[1]]
lda_accuracy

lda_error <- 1-lda_accuracy
lda_error

lda_sens <- sensitivity(c_mat)
lda_sens

lda_spec <- specificity(c_mat)
lda_spec
```
The accuracy of our prediction is 97.66% which gives us $\textbf{error rate}$ of 2.33%. The sensitivity is 95.80% and specificity is 100%.

```{r}

QDA.fit <- qda(Class~., data=bank.N)
QDA.pred <- predict(QDA.fit, bank.N)
summary(QDA.pred)

```
Summary of the QDA predict is shown above

```{r}
c_mat = table(QDA.pred$class, bank.N$Class)
c_mat

```
The confusion matrix of shows high false negative value. The model did great predction on 1s. 

```{r}
qda_accuracy <- confusionMatrix(c_mat)$overall[[1]]
as.matrix(c_mat)
qda_accuracy

qda_error <- 1-qda_accuracy
qda_error

qda_sens <- sensitivity(c_mat)
qda_sens

qda_spec <- specificity(c_mat)
qda_spec
```
The accuracy of our prediction is 98.54% which gives us $\textbf{error rate}$ of 0.14%. The sensitivity is 97.37% and specificity is 100%.


#k-NN (package: Class)

Using `knn`, fit k-NN classifiers for the entire dataset and calculate confusion matrix, (training) error rate, sensitivity/specificity for  k = 1 : 10. Compare them to corresponding results from logistic regression, LDA and QDA. Describe results of this comparison.  Discuss your low training  error rate for k-NN at k=1!

$\textbf{K-Nearest Neighbour(k-NN)}$

```{r}
bank.N_tr.tor <- bank.N[,5]
bank.N_tr <- scale(bank.N[,c(1,2,3,4)])

accuracy <- c()
for (n.K in c(1:10)) {
  #knn fit for the given data
  knn.fit <- knn(train=bank.N_tr, test=bank.N_tr, cl=bank.N_tr.tor, k=n.K)
  
  c_mat <- confusionMatrix(bank.N_tr.tor, knn.fit)
  if(n.K == 1){
    c_mat_k.1 <- table(bank.N_tr.tor, knn.fit)
  }

  print(paste("Value of K:", n.K))
  print(as.matrix(c_mat))
  accuracy <- append(accuracy, (c_mat$overall[[1]] * 100))
}
Error = 100 - accuracy
plot(Error, type = 'b', main = "Error plot for KNN", pch = 16,col = 20, xlab = "k")

```

The above output shows that the error rate increases with the increase in K. Without scaling the data, the model show accuracy of 100% for any k which says that the data might be overfitted.
KNN where k=1 gives us 100% accurate result. 

#TEST errors 

Now compare of logistic regression, LDA, QDA and k-NN. Using re-sampling approach, obtain test error as well as sensitivity and speciï¬city for each of these methods (logistic regression, LDA, QDA, k-NN with k = 1, 3, 5, 11, 21, 51, 101 ). Present your results in the form of BoxPlots, compare test error, sensitivity, and specificity for all these methods and discuss their relative performance.


$\textbf{Comparision of logistic regression, LDA, QDA and k-NN}$
Comparing the error rate, sensitivity and specificity of the test data.

```{r}
ntry <- 7

logr_error_rate <- NULL
qda_error_rate <- NULL
lda_error_rate <- NULL
knn_error_rate <- NULL


logr_sens_rate <- NULL
qda_sens_rate <- NULL
lda_sens_rate <- NULL
knn_sens_rate <- NULL

logr_spec_rate <- NULL
qda_spec_rate <- NULL
lda_spec_rate <- NULL
knn_spec_rate <-NULL


for(i in 1:ntry)
{
tr_idx  <- sample(nrow(bank.N),round(0.8*nrow(bank.N),digits=0)); length(tr_idx)
ts_idx  <- -tr_idx
bank.N_tr <- bank.N[tr_idx, ]; dim(bank.N_tr)
bank.N_ts <- bank.N[ts_idx, ]; dim(bank.N_ts)
bank.N_tr_f.tor <- bank.N[tr_idx, "Class"]
bank.N_ts_f.tor <- bank.N[ts_idx, "Class"]

# MODELS

#Logistic Regression Model
L.R.fit = glm(formula = Class ~ ., family = "binomial", data = bank.N_tr)
L.R.prob <- predict(L.R.fit,newdata=bank.N_ts,type="response")
L.R.response <- predict(L.R.fit,newdata=bank.N_ts,type="response")

conf_matrix = table(bank.N_ts[,"Class"], L.R.response > 0.5)
#calculate the error rate, sensitivity and specificity for logistic regresstion

logr_test_accuracy = sum(diag(conf_matrix))/sum(conf_matrix)
logr_test_error = 1 - logr_test_accuracy
logr_test_sensitivity = conf_matrix[2,2]/sum(conf_matrix[2,])
logr_test_specificity = conf_matrix[1,1]/sum(conf_matrix[1,])

logr_error_rate <- rbind(logr_error_rate,logr_test_error)
logr_sens_rate <- rbind(logr_sens_rate,logr_test_sensitivity)
logr_spec_rate <- rbind(logr_spec_rate,logr_test_specificity)

#LDA Model
LDA.fit <- lda(Class~., data=bank.N_tr)
LDA.pred <- predict(LDA.fit, bank.N_ts)

c_mat <- table(LDA.pred$class, bank.N_ts$Class)

#calculate the error rate, sensitivity and specificity for LDA
LDA_test_accuracy<-accuracy <- confusionMatrix(c_mat)$overall[[1]]
LDA_test_error <- 1-LDA_test_accuracy
LDA_test_sensitivity <- sensitivity(c_mat)
LDA_test_specificity <- specificity(c_mat)

lda_error_rate <- rbind(lda_error_rate,LDA_test_error)
lda_sens_rate <- rbind(lda_sens_rate,LDA_test_sensitivity)
lda_spec_rate <- rbind(lda_spec_rate,LDA_test_specificity)

#QDA Model
QDA.fit <- qda(Class~., data=bank.N_tr)
QDA.pred <- predict(QDA.fit, bank.N_ts)

c_mat = table(QDA.pred$class, bank.N_ts$Class)

#calculate the error rate, sensitivity and specificity for QDA
QDA_test_accuracy<-accuracy <- confusionMatrix(c_mat)$overall[[1]]
QDA_test_error <- 1-QDA_test_accuracy
QDA_test_sensitivity <- sensitivity(c_mat)
QDA_test_specificity <- specificity(c_mat)

qda_error_rate <- rbind(qda_error_rate,QDA_test_error)
qda_sens_rate <- rbind(qda_sens_rate,QDA_test_sensitivity)
qda_spec_rate <- rbind(qda_spec_rate,QDA_test_specificity)
}
```

The following is the test error rate for KNN 
```{r}
accuracy <- c()
for (n.K in c(1, 3, 5, 11, 21, 51, 101)) {
  #knn fit for the given data
  knn.fit <- knn(train=bank.N_tr, test=bank.N_ts, cl=bank.N_tr_f.tor, k=n.K)
  
  conf_matrix <- table(bank.N_ts_f.tor, knn.fit)

  knn_test_accuracy = sum(diag(conf_matrix))/sum(conf_matrix)
  knn_test_accuracy

  knn_test_error = 1 - knn_test_accuracy
  knn_test_error

  knn_test_sensitivity = conf_matrix[2,2]/sum(conf_matrix[2,])
  knn_test_sensitivity

  knn_test_specificity = conf_matrix[1,1]/sum(conf_matrix[1,])
  knn_test_specificity

###
  knn_error_rate <- rbind(knn_error_rate,knn_test_error)
  knn_sens_rate <- rbind(knn_sens_rate,knn_test_sensitivity)
  knn_spec_rate <- rbind(knn_spec_rate,knn_test_specificity)

}

```

$\textbf{Comparision table}$ for error rate, sensitivity and specificity of the test. 
```{r}

error <- data.frame(logr_error_rate = logr_error_rate[,1] * 100, lda_error_rate = lda_error_rate[,1] * 100, qda_error_rate = qda_error_rate[,1]*100, knn_error_rate = knn_error_rate[,1]*100)
sensitivity <- data.frame(logr_sens_rate = logr_sens_rate[,1] * 100, lda_sens_rate = lda_sens_rate[,1] * 100, qda_sens_rate = qda_sens_rate[,1]*100, knn_sens_rate = knn_sens_rate[,1]*100)
specificity <- data.frame(logr_spec_rate = logr_spec_rate[,1] * 100, lda_spec_rate = lda_spec_rate[,1] * 100, qda_spec_rate = qda_spec_rate[,1]*100, knn_spec_rate = knn_spec_rate[,1]*100)

error
sensitivity
specificity

```
$\textbf{Box plot comparision}$
Comparision of Logistic Regresssion, LDA, QDA and K-NN test result is show in box plot below

```{r}
ggplot(stack(error), aes(x = ind, y = values, fill = ind)) +
  geom_boxplot() +
  ylab("Error") +
  xlab("Performance")


ggplot(stack(sensitivity), aes(x = ind, y = values, fill = ind)) +
  geom_boxplot() +
  ylab("Sensitivity")+
  xlab("Performance")

ggplot(stack(specificity), aes(x = ind, y = values, fill = ind)) +
  geom_boxplot() +
  ylab("Specificity")+
  xlab("Performance")



```

The above plot show that KNN did a great predition. K = 1 in KNN model gives us 100% accurate result. 
The sensitivity plot shows the percentage of the bank note that is correctly identified as having the condition. KNN and logistic regression gives us high rate of sensitivity.
The specificity plot shows that KNN has high variance of the specificity. While lda and qda has high specificity rate which concludes that lda and qda did a great job in identifying the percentage of bank note not having the condition. 

#ROC and AUC

Plot the Region of Convergence "ROC" curve and find the numerical value of Area Under Curve "AUC"
$\textbf{Region of Convergence and Area Under the curve}$
Region of convergence for entire dataset
```{r}
predicted_prob<-predict(L.R.fit,type="response")
roccurve <- roc(L.R.fit$y, predicted_prob)
roccurve
plot(roccurve)
```

Manually calculation region of convergence for the test data
```{r}
L.R.prob <- predict(L.R.fit,newdata=bank.N_ts,type="response")
roc.pred <- prediction(L.R.prob,bank.N_ts_f.tor)
roc.tpr.fpr <- performance(roc.pred,measure="tpr", x.measure = "fpr")
plot(roc.tpr.fpr, col="blue3",col.lab="red3",
     xlab="False Positive(FP), rate",
     ylab="True Positive(TP), rate")

roc.auc    <- performance(roc.pred,measure="auc")
Area_Under <- roc.auc@y.values[[1]]
Area_Under
```
The area under the curve is 0.99.
